\begin{acknowledgements}      

Though only my name appears on the cover of this thesis, a great many people have contributed to its production. I owe my gratitude to all those people who have made this thesis possible and because of whom my graduate experience has been one that I will cherish forever.
My deepest gratitude is to my supervisor, Prof. Nguyen Khanh Van. I have been amazingly fortunate to have an supervisor who gave me the freedom to explore on my own, and at the same time the guidance to recover when my steps faltered. He taught me how to question thoughts and express ideas.

Mrs. Nguyen Phi Le's insightful comments and constructive criticisms at different stages of my research were thought-provoking and they helped me focus my ideas. I am also thankful to her for reading my reports, commenting on my views and helping me understand and enrich my ideas.

I am also indebted to the members of SEDIC Lab with whom I have interacted during the course of my studies. Particularly, I would like to acknowledge Mr. Nguyen Duc Trong, Mr. Truong Thao Nguyen for the many valuable discussions that helped me understand my research area better.

I am also grateful to the current lecturers at School of Information and Communication Technology, for their various forms of support during my college study.

Many friends have helped me stay sane through these difficult years -- Đô, Lù, and Na and especially \#Reiju Vinsmoke. Their support and care helped me overcome setbacks and stay focused on my study. Thank you for giving me hope and bringing light to my life.

Most importantly, none of this would have been possible without the love and patience of my family. My immediate family to whom this thesis is dedicated to, has been a constant source of love, concern, support and strength all these years. I would like to express my heart-felt gratitude to my family.

\end{acknowledgements}
  